Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-03-29 paper.tex
test
updated figures+Theory

%NEW
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{{\rm ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{{\rmn vir}}
\newcommand{\gal}{{\rmn gal}}
\newcommand{\iruv}{{\rmn sd}}
\newcommand{\clu}{{\rmn cl}}
\newcommand{\sfe}{{\rmn sfe}}
\newcommand{\sub}{{\rmn sub}}
\newcommand{\msun}{{M_\odot}}

%OLD
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\rmn}{\mathrm}
\newcommand{\GeV}{\,{\rm GeV}}
\newcommand{\TeV}{\,{\rm TeV}}
\newcommand{\eV}{\,{\rm eV}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\mx}{\ensuremath{m_{\chi}}}
\newcommand{\ga}{\gamma}
\newcommand{\ngamma}{\ensuremath{N_{\gamma}}}
\newcommand{\ngammaj}{\ensuremath{N_{\gamma,j}}}
\newcommand{\sigmaannv}{\ensuremath{\langle\sigma v\rangle}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}
\newcommand{\ic}{\mathrm{IC}}
\newcommand{\egamma}{\ensuremath{E_{\gamma}}}
\newcommand{\dic}{\mathrm{dIC}}
\newcommand{\fsr}{\mathrm{FSR}}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\rhos}{\ensuremath{\rho_s}}
\newcommand{\rs}{\ensuremath{r_s}}
\newcommand{\rvir}{r_{200}}
\newcommand{\mvir}{M_{200}}
\newcommand{\rhoc}{\ensuremath{\rho_c}}
\newcommand{\B}{\rmn{B}}
\newcommand{\e}{\rmn{e}}
\newcommand{\eg}{E_\gamma}
\newcommand{\ir}{\rmn{ir}}
\newcommand{\Km}{\rmn{K}`}
\newcommand{\Imm}{\rmn{I}`}
\newcommand{\Jm}{\rmn{J}^*}
\newcommand{\Jmm}{\rmn{J}`}

@@

include contributions from four main processes; leptophilic DM

@@

where $ q_{\pi^0-\gamma}(\eg,r)$ denotes the CR to gamma-ray source
function (see Sect.~\ref{sect:CRs} and \cite{2010MNRAS.409..449P}
for more details). The total differerntial boost factor for DM is the
product of enhancement factors from SFE $\B_\sfe(\sigv)$ (see
Sect.~\ref{sect:LP}) and from substructure enhancement over the
smooth halo contribution $\B_{\sub,i}(r) = 1+q_\sub(r)/q_{\rmn{sm},i}(r)$

@@

\quad q_{\rmn{sm},i} (\egamma,r) = \sum_j

@@

We represent the annihilation rate density with $\Gamma_j =
(\rho/\mx)^2 \, \sigmaannv_j/2$, where $j$ runs over all $\gamma$-ray

@@

  \ngammaj}{\dd\eg}$ and annihilation cross-section $\sigmaannv_j$.
The DM density is modeled by an Einasto density profile

@@

@@

that is slightly shallower than the conventional Navaro-Frenk-White
(NFW) profile in the center, but provide a better fit to simulated DM
halos \cite{2010MNRAS.402...21N}. We denote the density where the
profile has a slope of $-2$ by $\rho_{-2}$, and the radius by
$r_{-2}$. Assuming that all the flux from an NFW profile originate
$\Delta=200$ times the critical density $\rhoc$ of the universe. The
characteristic density of the NFW profile $\rhos=(\Delta\rhoc/3)\,

@@

according to \cite{2003ApJ...584..702H}.

\subsubsection{Substructures}
\label{sect:subst}
High-resolution DM simulations of the Milky Way (MW) find vast amount
of substrucures in the periphery of DM halos. Since the rate of which
DM is annihilating depends on the density squared, the resulting flux
is boosted compared to the smooth density distribution. Recent
simulations suggest a boost from substructures of the order of ten up
to a few 100 \cite{2008MNRAS.391.1685S, 2008JPhCS.125a2008K} for a MW
size halo. Based on the result from \cite{2008Natur.456...73S}, we
find the best fit\footnote{Our approach of fitting the scaling
  behavior of $\rmn{L_{sub}}(<r)$ directly from numerical simulations
  self-consistently accounts for the radial dependence of the
  substructure concentration \cite{2008MNRAS.391.1685S}.} to the
luminosity from smooth substructures (i.e. substructures within
substructures are not included) inside radius $r$ with
\begin{equation}
  {\rm L_{sub}}(<r)= 0.76\,C_1\,C_2(\mvir)\,{\rm L_{200sm}}\,x^{0.95
    \displaystyle{x}^{-0.27}}\,,
  \label{eq:LSub}
\end{equation}
where ${\rm L_{200sm}}$ is the luminosity from the smooth halo without
substructures within $\rvir$ and $x=
\left(\frac{r}{\rvir}\right)$. The first normalization constant is
given by
\begin{equation}
  C_1=\left(\frac{M_{\rm res,sim}}{M_{\rm lim}}\right)^{0.226}\,,
   \label{eq:C1}
\end{equation}
where $M_{\rm res,sim}=10^5\msun$ is the minimum subhalo mass in the
MW simulation, and $M_{\rm lim}$ the free streaming mass. Here we
assume that the power-law scaling relation in Eq.~(\ref{eq:C1}) is
valid down to the minimum mass of DM halos, which is conventionaly in
the cold dark matter universe
\cite{2001PhRvD..64h3507H,2005JCAP...08..003G} taken to be
$10^{-6}\msun$. Note that potentially the power-law could flatten
torwards smaller mass scales although current simulations show no hits
of such a behaviour and in addition we are approaching the asymptotic
behavior in the power spectrum on these scales(CHECK). For DM halos
more (less) massive than the MW we expect a larger (smaller) boost
from substructures, simply because of the larger (smaller) mass range
down to the minimum mass $M_\rmn{lim}$. We capture this halo mass
dependence with the second normalization in Eq.~(\ref{eq:LSub}):
\begin{equation}
C_2(\mvir)=\left(\frac{\mvir}{M_{\rm 200sim}}\right)^{0.226}\,,
 \label{eq:C2}
\end{equation}
where $M_{\rm 200,sim}=1.9\times10^{12}\msun$ is the mass of the MW
halo in the simulation \cite{2008Natur.456...73S}.

large uncertaint observationally in density profiles

In Fig.~\ref{fig:radial_lum} we compare the radial dependence of the
accumulative luminosity from different smooth cluster density profiles
to the substructure boosted luminosity at different mass scles. The

\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/dens.prof.eps}
\caption{\it Radial dependence of various luminosity
  contributions. The solid lines show the accumulative smooth
  luminosity from a cluster with the mass $\mvir=10^{15}\,\msun$ for
  three different density profiles; an Einasto profile with
  $\alpha=0.17$ (light blue), a cuspy NFW profile with $\beta=1.0$
  (dark blue), and a core NFW profile with $\beta=0.6$ (purple). The
  dashed lines show the accumulative luminosity from substructures for
  three different mass scales; an $\mvir=10^{15}\,\msun$ galaxy
  cluster (red), an $\mvir=10^{12}\,\msun$ galaxy (orange), and an
  $\mvir=10^{8}\,\msun$ dwarf galaxy (green). All luminosities have
  been normalized with the luminosity within $\rvir$ from a cuspy NFW
  profile. Note the large expected boost from substructures in
  clusters ($\sim1000$), and the relative small boost in dwarf
  galaxies ($\sim20$).}
 \label{fig:radial_lum}
\end{figure}

\begin{figure}
  \includegraphics[width=0.99\columnwidth]{figures/emissiv.sub.eps}
  \caption{\it Volume weighted luminosity for different at different
    scales. The solid lines show the emissivity from a NFW density
    profile for three different mass scales; a $\mvir=10^{15}\,\msun$
    galaxy cluster (red), a $\mvir=10^{12}\,\msun$ galaxy (orange), a
    $\mvir=10^{8}\,\msun$ dwarf galaxy (green). The dashed lines show
    the emissivity from substructures for the same three mass
    scales. All emissivities have been normalized with the luminosity
    from a NFW profile within $\rvir$.}
  \label{fig2}
\end{figure}

\subsubsection{Cosmic-ray induced gamma-ray emission}
\label{sect:CRs}
The formation process of a galaxy cluster is a very energetic
processes that induces both turbulence as well as frequently occurring
merging and accretion shocks. These processes are thought to
accelerate large quantities of relativistic non-thermal electrons and
protons to high energies. On smaller scales, such as galaxies, there
is convincing evidence of such non-thermal populations. Especially, in
the MW, the cosmic rays are observed directly as well as indirectly
through radio, X-ray, and gamma-ray emission. On larger scales of the
order of few 100 kpc up to Mpc, there are a vast number of
observations of radio emission coming from radio mini halos in the
center of cooling flow clusters, radio relics in the periphery of
clusters \cite{2004rcfg.proc..335K} as well smooth giant radio halos
that have been observed in more than 50 clusters
\cite{2003ASPC..301..143F,2008SSRv..134...93F}. Most of this observed
radio emission is expected to emerge from synchrotron emitting
cosmic-ray electrons whose precise origin in especially relics and
giant radio halos is still not settled. However, due to the smoothness
of the extended emission and the observed power-law slope it seems
quite natural that they are of hadronic origin. Especilly since the
long cooling time of cosmic ray protons (CRs) allows for both an
extended and large population of CRs to build up over the cluster
history \cite{1997ApJ...487..529B}.

When the CRs interact with ambient gas protons, it results in the
production of both charged and neutral pions that decay into
electrons, neutrinos, and gamma-rays. The production of these
secondaries depend both on the gas and CR densities in the cluster,
where the CR density roughly trace the gas outside the core regime and
is slightly enhance in the center. This density scaling implies that
clusters are great targets for Cherenkov telescopes with a high
sensitivity for the central parts of close by clusters. Detecting the
cluster gamma-ray emission is crucial in this respect as it
potentially provides the unique and unambiguous evidence of CR
populations in clusters through observing the $\pi^0$ bump at about
100~MeV in the spectra. We adopt the spectrally and spatially
universal gamma-ray model developed by Pinzke \& Pfrommer
\cite{2010MNRAS.409..449P} to estimate the emission from decaying
$\pi^0$:s that dominates over the inverse Compton emission from
primary and secondary electrons above 100~MeV in clusters. The
gamma-ray formalism was derived from high resolution simulations of
clusters of galaxies that included radiative hydrodynamics, star
formation, supernova feedback, and followed CR physics using a novel
formulation that trace the most important injection and loss processes
self-consistently while accounting for the CR pressure in the equation
of motion
\cite{2008A&A...481...33J,2007A&A...473...41E,2006MNRAS.367..113P}.
We note that the overall normalization of the CR and gamma-ray
distribution scales with the maximum acceleration efficiency at
structure formation shock waves. Following recent observations at
supernova remnants \cite{2009Sci...325..719H} as well as theoretical
studies \cite{2005ApJ...620...44K}, we adopt a realistic value of this
parameter and assume that 50\% of the dissipated energy at strong
shocks is injected into CRs while this efficiency rapidly decreases
for weaker shocks \cite{2007A&A...473...41E}.

\subsection{Particle physics}

@@

\subsubsection{Supersymetric dark matter}
neutralino, benchmark models, continuum emission, electron spectra

@@

\subsection{Radiative processes}

\subsubsection{Inverse Compton}
The source function of inverse Compton emission resulting from DM
annihilating is given by
\begin{equation}
  q_{\rm IC}\left(E_\gamma, r\right) = \int
  \dd E_\e\, \frac{\dd n_\e}{\dd E_\e} 
  P_{\rm IC}\left(E_\gamma,E_\e\right) ,
  \label{eq:ICemiss}
\end{equation}
where $P_{\rm IC}$ is derived by convolving the IC cross-section with
the differential target photon number density:
\begin{eqnarray}
\label{eq:KN_spec}
P_\rmn{IC} &=&
\frac{3\sigma_\rmn{T}m^2_\e c^5}{4E_\e^2}
\int\frac{n_\rmn{ph}(\eph)\dd \eph}{\eph}\nonumber \\
&\times&\left[2q
  \ln{q}+(1+2q)(1-q)+ \frac{1}{2}\frac{\left(\Gamma_\e q\right)^2}
     {1+\Gamma_\e q}\left(1-q\right)\right]\,,\nonumber \\
&&
\end{eqnarray}
where
\begin{equation}
\Gamma_\e=\frac{4\eph E_\e}{\left(m_\e c^2\right)^2}\,,\qquad \rmn{and} \qquad  
q=\frac{\eg}{\Gamma_\e(E_\e-\eg)}\,.
\end{equation}
We account for two major contributions to the radiative background
field $n_\ph$; the IR and UV light coming from emitting dust and
starlight where $n_\ph\equiv\frac{\dd^2 N_\ph^\clu}{\dd V \,\dd
  \eph}(\eph,r)= \frac{u_\iruv^\clu(\eph,r)}{\eph^2}$ and
$u_\iruv^\clu(\eph,r)$ is given by Eq.~(\ref{eq:E_IR_UV}). We model the
CMB photon spectrum as a photon gas that is isotropically distributed and
follows a black body spectrum with temperature $T$:
\begin{equation}
\label{eq:photon_gas}
  n_\rmn{ph}(\eph) = \frac{\dd^2 N_\ph}{\dd V \,\dd \eph} =
  \frac{1}{\pi^2(\hbar c)^3}\frac{\eph^2}{\exp(\eph/k_\B T)-1}\,.
\end{equation}
Here the the typical energy of a photon before scattering is given by
$<E_\ph>=\epsilon_\ph/n_\ph\approx\,2.70 k_\B T$, where $n_\ph$ and
$\epsilon_\ph$ are the number- and energy-density derived by
integrating $n_\ph(E_\ph)$ and $E_\ph n_\ph(\eph)$ over the photon
energy $\eph$, respectively.

The equilibrium spectrum of the electrons plus positrons in
Eq.~(\ref{eq:ICemiss}) is given by $\frac{\dd n_\e}{\dd E_\e}$ and
derived from the CRe diffusion equation:
\begin{eqnarray}
\label{eq:CRdiffusion}
\frac{\partial}{\partial t}\left(\frac{\dd n_\e}{\dd E_\e}\right) = 
&\nabla& \left[D\left(E_\e,\bx\right)\nabla\frac{\dd n_\e}{\dd E_\e}\right] + 
\frac{\partial}{\partial E_\e}
\left[b\left(E_\e,\bx\right) \frac{\dd n_\e}{\dd E_\e}\right]
 \nonumber \\
&+& Q(E_\e,\bx)\,,
\end{eqnarray}
where $D(E_\e,\bx)$ the diffusion coefficient and $b(E_\e,\bx)$ the
electron energy loss term. The source function $Q(E_\e,\bx)$ shows
the number of particles produced per unit time, energy and volume
element that are produced at the position $\bx$:
\begin{equation}
Q(E_\e,r)=\sum_f\frac{\dd N_\e^f}{\dd E_\e}(E_\e) B_f \Gamma(r)\,.
\end{equation}
The subscript $f$ ...(see cola), and $B_f$ is the branching factor.
Here $\frac{\dd N_\e^f}{\dd E_\e}$ denote the differential number of
electrons plus positrons resulting from an annihilation event. For
neutralinos annihilating directly into an electron and positron we
model the differential spectrum with $\frac{\dd N_\e^f}{\dd E_\e}=
2\delta(E_\e-m_\chi c^2)$. We use DarkSUSY to compute both the spectra
when neutralinos annihilate directly into a $\mu^+$ and $\mu^-$ in
the leptophilic model as well as in the four BM models where a
fraction of the annihilating neutralinos is converted into electrons
and positrons (see \cite{2009JCAP...01..016B} and references therein,
also see fit).

>>>>

%NEW
\newcommand{\rmn}{\mathrm}
\newcommand{\bx}{{\bf x}}
\newcommand{\fg}{{F_\gamma}}
\newcommand{\sg}{{S_\gamma}}
\newcommand{\psf}{\theta_\rmn{res}}
\newcommand{\ph}{\rmn{ph}}
\newcommand{\eph}{E_\ph}
\newcommand{\nph}{N_\ph}
\newcommand{\vir}{\rmn{vir}}
\newcommand{\gal}{\rmn{gal}}
\newcommand{\sd}{\rmn{SD}}
\newcommand{\clu}{\rmn{cl}}
\newcommand{\sfe}{\rmn{sfe}}
\newcommand{\sub}{\rmn{sub}}
\newcommand{\msun}{M_\odot}
\newcommand{\ee}{E_\rmn{e}}
\newcommand{\stars}{\rmn{stars}}
\newcommand{\dust}{\rmn{dust}}
\newcommand{\lx}{L_\rmn{X}}
\newcommand{\s}{\rmn{s}}

%OLD
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\gev}{\rm{GeV}}
\newcommand{\tev}{\rm{TeV}}
\newcommand{\ev}{\rm{eV}}
\newcommand{\dd}{\rmn{d}}
\newcommand{\mx}{\ensuremath{m_{\chi}}}
\newcommand{\ga}{\gamma}
\newcommand{\ngamma}{\ensuremath{N_{\gamma}}}
\newcommand{\ngammaj}{\ensuremath{N_{\gamma,j}}}
\newcommand{\sigmaannv}{\ensuremath{\langle\sigma v\rangle}}
\newcommand{\sigv}{\ensuremath{\sigma_v}}
\newcommand{\ic}{\rmn{IC}}
\newcommand{\egamma}{\ensuremath{E_{\gamma}}}
\newcommand{\dic}{\rmn{dIC}}
\newcommand{\fsr}{\rmn{FSR}}
\newcommand{\CR}{\rmn{CR}}
\newcommand{\rhos}{\ensuremath{\rho_s}}
\newcommand{\rs}{\ensuremath{r_s}}
\newcommand{\rvir}{r_{200}}
\newcommand{\mvir}{M_{200}}
\newcommand{\rhoc}{\ensuremath{\rho_c}}
\newcommand{\B}{\rmn{B}}
\newcommand{\e}{\rmn{e}}
\newcommand{\eg}{E_\gamma}
\newcommand{\ir}{\rmn{ir}}
\newcommand{\Km}{\rmn{K}`}
\newcommand{\Imm}{\rmn{I}`}
\newcommand{\Jm}{\rmn{J}^*}
\newcommand{\Jmm}{\rmn{J}`}

@@

include contributions from five main processes; leptophilic DM

@@

where the subscript $i$ runs over the gamma-ray producing DM
channels. The differential CR to gamma-ray source function is denoted
by $q_{\pi^0-\gamma}(\eg,r)$, where we refer the reader to
Sect.~\ref{sect:CRs} and \cite{2010MNRAS.409..449P} for further
details. The total differential boost factor for DM is the product of
enhancement factors from SFE $\B_\sfe(\sigv)$ (see
Sect.~\ref{sect:LP}) and from substructure enhancement over the smooth
halo contribution $\B_{\sub,i}(r) = 1+q_\sub(r)/q_{\rmn{sm},i}(r)$

@@

\label{eq:q_sm}
q_{\rmn{sm},i} (\egamma,r) = \sum_j

@@

Here the annihilation rate density is given by 
\label{eq:ann_rate}
\Gamma_j(r) = \frac{1}{2} \left(\frac{\rho(r)}{\mx}\right)^2 
\, \sigmaannv_j\,
where the subscript $j$ runs over all kinematically allowed gamma-ray

@@

  \ngammaj}{\dd\eg}$ and annihilation cross-section $\sigmaannv_j$. We
denote the DM mass with $\mx$ and the DM density profile with
$\rho(r)$. Typically the universal Navarro-Frenk-White (NFW) density
profile provide a good fit to both the observed and simulated
clusters. It can be considered as a special case of the 5-parameter profile:

@@

\rho(r) = \frac{\rho_\s}{\left(r/r_\s+\right)^\gamma
  \left[1+\left(r/r_\s\right)^\alpha\right]^\delta}\,,\quad
\delta=\frac{\beta-\gamma}{\alpha}\,,
\label{eq:rho_nfw}

@@

where a cuspy NFW profile is given by
$(\alpha,\beta,\gamma)=(1,3,1)$. The scaling radius is given by $r_\s$
and the characteristic density by $\rhos=(\Delta\rhoc/3)\,

@@

according to \cite{2003ApJ...584..702H}. In this work we choose to
model the DM density by an Einasto density profile
that is slightly shallower in the center than the conventional
Navarro-Frenk-White (NFW) profile, but provide a better fit to
simulated DM halos \cite{2010MNRAS.402...21N}. We denote the density
where the profile has a slope of $-2$ by $\rho_{-2}$, and the radius
by $r_{-2}$. Assuming that all the flux from an NFW profile originate
$\Delta=200$ times the critical density $\rhoc$ of the universe.

\subsubsection{Substructures}
\label{sect:subst}
High-resolution DM only simulations of Milky Way (MW) type halos find
substantial amount of substructures in the periphery of DM halos,
while the substructures in the center are erased due to dynamical
disruption. Since the rate of which DM is annihilating depends on the
density squared, the resulting flux is boosted compared to the smooth
density distribution. Recent simulations suggest a flux enhancement
from substructures of the order of ten up to a few 100
\cite{2008MNRAS.391.1685S, 2008JPhCS.125a2008K} for a MW size
halo. Using the Aquarius simulations \cite{2008Natur.456...73S}, we
fit \footnote{Our approach of fitting the scaling behavior of
  $\rmn{L_{sub}}(<r)$ directly from numerical simulations
  self-consistently accounts for the radial dependence of the
  substructure concentration \cite{2008MNRAS.391.1685S}.} to the
luminosity from smooth substructures (i.e. substructures within
substructures are not included) inside radius $r$. The best fit is
given by
\begin{equation}
  {\rm L_{sub}}(<r)= 0.76\,C_1\,C_2(\mvir)\,{\rm L_{200sm}}\,x^{0.95
    \displaystyle{x}^{-0.27}}\,,
  \label{eq:LSub}
\end{equation}
where ${\rm L_{200sm}}$ is the luminosity from the smooth halo without
substructures within $\rvir$ and $x=
\left(\frac{r}{\rvir}\right)$. The first normalization constant is
derived from the simulations:
\begin{equation}
  C_1=\left(\frac{M_{\rm res,sim}}{M_{\rm lim}}\right)^{0.226}\,,
   \label{eq:C1}
\end{equation}
where $M_{\rm res,sim}=10^5\msun$ is the minimum subhalo mass in the
MW simulation, and $M_{\rm lim}$ the free streaming mass. Here we
assume that the power-law scaling relation in Eq.~(\ref{eq:C1}) is
valid down to the minimum mass of DM halos, which is conventionally in
the cold dark matter universe
\cite{2001PhRvD..64h3507H,2005JCAP...08..003G} taken to be
$10^{-6}\msun$. Note that potentially the power-law could flatten
towards smaller mass scales although current simulations show no hits
of such a behavior and in addition we are approaching the asymptotic
behavior in the power spectrum on these scales(CHECK). For DM halos
more (less) massive than the MW we expect a larger (smaller) boost
from substructures, simply because of the larger (smaller) mass range
down to the minimum mass $M_\rmn{lim}$. We capture this halo mass
dependence with the second normalization in Eq.~(\ref{eq:LSub}):
\begin{equation}
C_2(\mvir)=\left(\frac{\mvir}{M_{\rm 200sim}}\right)^{0.226}\,,
 \label{eq:C2}
\end{equation}
where $M_{\rm 200sim}=1.9\times10^{12}\msun$ is the mass of the MW
halo in the simulation \cite{2008Natur.456...73S}.

Recent detailed galaxy cluster observations find a core like density
profile with an inner slope of $\beta=0.6$, where $\beta\gtrsim 1$ can
be ruled out to a $>95$\% confidence \cite{2011ApJ...728L..39N}. This
finding is hard to reconcile with simulations that typically prefere a
steeper central profile. The different density profiles have a large
impact on the luminosity from annihilating DM, although the details of
the density profile can be neglected compared to the dominating boost
from substructures. In Fig.~\ref{fig:radial_lum} we compare the radial
dependence of the accumulative luminosity from different smooth
cluster density profiles to the substructure boosted luminosity at
different mass scales. We recalculate the density contrast for the
core like density profile $\rhos$ and rescale the concentration
parameter in Eq.~(\ref{eq:cfit}) with $300/160$
\cite{2011ApJ...728L..39N} to account for the more centralized scale
radius in a core like cluster. The emission coming from the core
compared to a cuspy like density profile is about 30\% larger within
$\rvir$. This difference is build up within 10\% of $\rvir$
(i.e. close to $r_\rmn{s}$). The emission from an Einasto density
profile is about 50\% larger than the cuspy NFW profile in the
periphery of the cluster, where the difference is mainly build up at a
few percent of $\rvir$. The substructure boosted luminosities is
neglible in the center of halos, but integrated out to $\rvir$ the
boost is about; 10 for dwarf galaxies, $10^2$ for galaxies, and $10^3$
for galaxy clusters.
\begin{figure}%[t]
 \includegraphics[width=0.99\columnwidth]{figures/dens.prof.eps}
\caption{\it Radial dependence of various luminosity
  contributions. The solid lines show the accumulative smooth
  luminosity from a cluster with the mass $\mvir=10^{15}\,\msun$ for
  three different density profiles; an Einasto profile with
  $\alpha=0.17$ (light blue), a cuspy NFW profile with $\beta=1.0$
  (dark blue), and a core NFW profile with $\beta=0.6$ (purple). The
  dashed lines show the accumulative luminosity from substructures for
  three different mass scales; an $\mvir=10^{15}\,\msun$ galaxy
  cluster (red), an $\mvir=10^{12}\,\msun$ galaxy (orange), and an
  $\mvir=10^{8}\,\msun$ dwarf galaxy (green). All luminosities have
  been normalized with the luminosity within $\rvir$ from a cuspy NFW
  profile. We have assumed the standard value for the limiting
  substructure mass of $10^{-6}\,\msun$. Note the large expected
  boost from substructures in clusters ($\sim1000$), and the relative
  small boost in dwarf galaxies ($\sim20$).}
 \label{fig:radial_lum}
\end{figure}

Even though the bulk of substructures have been erased in the central
regions of DM halos, a cluster in projection has a significant
enhancement due to substructures at a radius of just a few percent of
$\rvir$. In Fig.~\ref{fig:radial_emis} we show the luminosity
dominating radius for substructures and smooth density profiles, both
for three different mass scales. It clearly shows how that the main
contribution from the substructures is coming from outer parts of the
DM halos, while the main contribution from the smooth density profiles
peaks around the scale radius. NOT QUITE, ALSO FIGURE IS SHOWING DIFF
VOLUME, FAC 3 DIFFERENCE, WHAT DO WE REALLY LEARN FROM THE FIGURE
\begin{figure}
  \includegraphics[width=0.99\columnwidth]{figures/emissiv.sub.eps}
  \caption{\it Radial dependence of differential luminosity at
    different mass scales. The solid lines show the emissivity from a
    cuspy $\beta=1.0$ NFW density profile for three different mass
    scales; an $\mvir=10^{15}\,\msun$ galaxy cluster (red), an
    $\mvir=10^{12}\,\msun$ galaxy (orange), and an
    $\mvir=10^{8}\,\msun$ dwarf galaxy (green). The dashed lines show
    the emissivity from substructures for the same three mass
    scales. All emissivities have been normalized with the luminosity
    within $\rvir$ from a cuspy NFW profile. We have assumed the
    standard value for the limiting substructure mass of
    $10^{-6}\,\msun$.}
  \label{fig:radial_emis}
\end{figure}

\subsubsection{Cosmic-ray induced gamma-ray emission}
\label{sect:CRs}
The formation process of a galaxy cluster is a very energetic
processes that induces both turbulence as well as frequently occurring
merging and accretion shocks. These processes are thought to
accelerate large quantities of relativistic non-thermal electrons and
protons to high energies. On smaller scales, such as galaxies, there
is convincing evidence of such non-thermal populations. Especially, in
the MW, the cosmic rays are observed directly as well as indirectly
through radio, X-ray, and gamma-ray emission. On larger scales of the
order of few 100 kpc up to Mpc, there are a vast number of
observations of radio emission coming from radio mini halos in the
center of cooling flow clusters, radio relics in the periphery of
clusters \cite{2004rcfg.proc..335K} as well smooth giant radio halos
that have been observed in more than 50 clusters
\cite{2003ASPC..301..143F,2008SSRv..134...93F}. Most of this observed
radio emission is expected to emerge from synchrotron emitting
cosmic-ray electrons whose precise origin in especially relics and
giant radio halos is still not settled. However, due to the smoothness
of the extended emission and the observed power-law slope it seems
quite natural that they are of hadronic origin. Especially since the
long cooling time of cosmic ray protons (CRs) allows for both an
extended and large population of CRs to build up over the cluster
history \cite{1997ApJ...487..529B}.

When the CRs interact with ambient gas protons, it results in the
production of both charged and neutral pions that decay into
electrons, neutrinos, and gamma-rays. The production of these
secondaries depend both on the gas and CR densities in the cluster,
where the CR density roughly trace the gas outside the core regime and
is slightly enhance in the center. This density scaling implies that
clusters are great targets for Cherenkov telescopes with a high
sensitivity for the central parts of close by clusters. Detecting the
cluster gamma-ray emission is crucial in this respect as it
potentially provides the unique and unambiguous evidence of CR
populations in clusters through observing the $\pi^0$ bump at about
100~MeV in the spectra.

We adopt the spectrally and spatially universal gamma-ray model
developed by Pinzke \& Pfrommer \cite{2010MNRAS.409..449P} to estimate
the emission from decaying $\pi^0$:s that dominates over the inverse
Compton emission from primary and secondary electrons above 100~MeV in
clusters. The gamma-ray formalism was derived from high resolution
simulations of clusters of galaxies that included radiative
hydrodynamics, star formation, supernova feedback, and followed CR
physics using a novel formulation that trace the most important
injection and loss processes self-consistently while accounting for
the CR pressure in the equation of motion
\cite{2008A&A...481...33J,2007A&A...473...41E,2006MNRAS.367..113P}.
We note that the overall normalization of the CR and gamma-ray
distribution scales with the maximum acceleration efficiency at
structure formation shock waves. Following recent observations at
supernova remnants \cite{2009Sci...325..719H} as well as theoretical
studies \cite{2005ApJ...620...44K}, we adopt a realistic value of this
parameter and assume that 50\% of the dissipated energy at strong
shocks is injected into CRs while this efficiency rapidly decreases
for weaker shocks \cite{2007A&A...473...41E}.

\subsection{Particle physics}
\label{sect:PF}

@@

\subsubsection{Supersymmetric dark matter}
neutralino, benchmark models, continuum emission, electron spectra

@@

\subsection{Radiative processes}
In this section we outline the basics of inverse Compton (IC)
emission. As target radiation fields we consider both cosmic microwave
background (CMB) photons, and the light from stars and dust (SD). We
derive a semi-analytic model from which we can estimate the SD
spectral and spatial distributions as a simple function of cluster
mass.

\subsubsection{Inverse Compton}
\label{sect:IC}
The standard IC source function is given by
\cite{1979rpa..book.....R}:
\begin{eqnarray}
  q_{\rmn IC}(\eg, r) &=&  \frac{\dd^3 N_\gamma}{\dd V\,\dd t\,\dd \eg} = 
 \frac{3}{4}c\,\sigma_\rmn{T}
\int\dd \eph \frac{n_\rmn{ph}(\eph)}{\eph}\nonumber\\
&\times& \int \dd \ee\,\frac{\dd n_\e}{\dd \ee}(\ee,r)\,
 \frac{\left(m_\e c^2\right)^2}{\ee^2}G(\Gamma_\e,q)\,,\nonumber\\
  \label{eq:ICemiss}
\end{eqnarray}
where $\ee$ is the energy of the upscattering electrons and $\eph$ is
the energy of the background photon field. We represent the Thomson
cross-section with $\sigma_\rmn{T}$ and the full differential
Klein-Nishina (KN) cross-section is captured by
\cite{1970RvMP...42..237B}:
\begin{equation}
\label{eq:KN_spec}
G(\Gamma_\e,q) = 2q\ln{q}+(1+2q)(1-q)+ 
\frac{1}{2}\frac{\left(\Gamma_\e q\right)^2\left(1-q\right)}
     {1+\Gamma_\e q}\,,
\end{equation}
where
\begin{equation}
\Gamma_\e=\frac{4\eph \ee}{\left(m_\e c^2\right)^2}\,,\qquad \rmn{and} \qquad  
q=\frac{\eg}{\Gamma_\e\left(\ee-\eg\right)}\,.
\end{equation}
The full KN cross-section accounts for the less efficient energy
transfer between the photon and electron and results in a steepening
of the IC gamma-ray spectrum. In the low energy Thomson regime the IC
spectrum is given by $E_\gamma^{-(\alpha_\e-1)/2}$, however when
$\Gamma_\e \gg 1$ the IC spectrum steepens due to the KN suppression
to $\eg^{-\alpha_\e}\log(\eg)$ .

We account for two major contributions to the number density of
radiative background fields $n_\ph$; the cosmic microwave background
(CMB) photons and the infra-red (IR) to ultra-violet (UV) light
emitted by starlight and dust (SD). The number density for the SD is
given by $n_\ph\equiv\frac{\dd^2 N_\ph}{\dd V \,\dd
  \eph}(\eph,r)= \frac{u_\sd(\eph,r)}{\eph^2}$ where the SD
energy density $u_\sd(\eph,r)$ is given in Eq.~(\ref{eq:U_SD}). We
model the CMB photon spectrum as a photon gas that is isotropically
distributed and follows a black body spectrum with the temperature
$T=2.73\,$~K:
\begin{equation}
\label{eq:photon_gas}
  n_\rmn{ph}(\eph) = \frac{\dd^2 N_\ph}{\dd V \,\dd \eph} =
  \frac{1}{\pi^2(\hbar c)^3}\frac{\eph^2}{\exp(\eph/k_\B T)-1}\,.
\end{equation}
Note that the typical energy of a black body photon before scattering
is given by $<\eph>=\epsilon_\ph/\tilde{n}_\ph\approx\,2.70 k_\B T$,
where $\tilde{n}_\ph$ and $\epsilon_\ph$ are the number- and
energy-density derived by integrating $n_\ph(\eph)$ and $\eph
n_\ph(\eph)$ over the photon energy $\eph$, respectively.
